---
title: "ppatil8_hw6"
author: "Prachi Vijay Patil"
date: "`r Sys.Date()`"
output: html_document
---

# 9.6 Lab 
# Support Vector Machine

# 9.6.1
```{r}
set.seed(1)
x <- matrix(rnorm(20 * 2), ncol = 2)
y <- c(rep(-1, 10), rep(1, 10))
x[y == 1, ] <- x[y == 1, ] + 1
plot(x, col = (3 - y))
dat <- data.frame(x = x, y = as.factor(y))
library(e1071)
svmfit <- svm(y ~ ., data = dat, kernel = "linear", cost = 10, scale = FALSE)
plot(svmfit, dat)
svmfit$index
summary(svmfit)
svmfit <- svm(y ~ ., data = dat, kernel = "linear", cost = 0.1, scale = FALSE)
plot(svmfit, dat)
svmfit$index

set.seed(1)
tune.out <- tune(svm, y ~ ., data = dat, kernel = "linear", 
                 ranges = list(cost = c(0.001, 0.01, 0.1, 1, 5, 10, 100)))
summary(tune.out)
bestmod <- tune.out$best.model
summary(bestmod)
xtest <- matrix(rnorm(20 * 2), ncol = 2)
ytest <- sample(c(-1, 1), 20, rep = TRUE)
xtest[ytest == 1, ] <- xtest[ytest == 1, ] + 1
testdat <- data.frame(x = xtest, y = as.factor(ytest))
ypred <- predict(bestmod, testdat)
table(predict = ypred, truth = testdat$y)
svmfit <- svm(y ~ ., data = dat, kernel = "linear", cost = 0.01, scale = FALSE)
ypred <- predict(svmfit, testdat)
table(predict = ypred, truth = testdat$y)
x[y == 1, ] <- x[y == 1, ] + 0.5
plot(x, col = (y + 5) / 2, pch = 19)
dat <- data.frame(x = x, y = as.factor(y))
svmfit <- svm(y ~ ., data = dat, kernel = "linear", cost = 1e5)
summary(svmfit)
plot(svmfit, dat)

svmfit <- svm(y ~., data = dat, kernel = "linear", cost = 1)
summary(svmfit)
plot(svmfit, dat)
```

# 9.6.2

```{r}
set.seed(1)
x <- matrix(rnorm(200 * 2), ncol = 2)
x[1:100, ] <- x[1:100, ] + 2
x[101:150, ] <- x[101:150, ] - 2
y <- c(rep(1, 150), rep(2, 50))
dat <- data.frame(x =x, y = as.factor(y))
plot(x, col = y)
train <- sample(200, 100)
svmfit <- svm(y ~ ., data = dat[train, ], kernel = "radial", gamma = 1, cost = 1)
plot(svmfit, dat[train, ])
summary(svmfit)
svmfit <- svm(y ~ ., data = dat[train, ], kernel = "radial", gamma = 1, cost = 1e5)
plot(svmfit, dat[train, ])
set.seed(1)
tune.out <- tune(svm, y ~ ., data = dat[train, ], 
                 kernel = "radial", 
                 ranges = list(
                   cost = c(0.1, 1, 10, 100, 1000),
                   gamma = c(0.5, 1, 2, 3, 4)
))
summary(tune.out)
table(
  true = dat[-train, "y"],
  pred = predict(
    tune.out$best.model, newdata = dat[-train, ]
  )
)
```

# 9.6.3

```{r}
library(ROCR)
rocplot <- function(pred, truth, ...) {
  predob <- prediction(pred, truth)
  perf <- performance(predob, "tpr", "fpr")
  plot(perf, ...)
}
svmfit.opt <- svm(y ~ ., data = dat[train, ],
                  kernel = "radial", gamma = 2, cost = 1,
                  decision.values = T)
fitted <- attributes(
  predict(svmfit.opt, dat[train, ], decision.values = TRUE)
)$decision.values
par(mfrow = c(1, 2))
rocplot(-fitted, dat[train, "y"], main = "Training Data")
svmfit.flex <- svm(y ~ ., data = dat[train, ], 
                   kernel = "radial", gamma = 50, cost = 1,
                   decision.values = T)
fitted <- attributes(
  predict(svmfit.flex, dat[train, ], decision.values = T)
)$decision.values
rocplot(-fitted, dat[train, "y"], add = T, col = "red")
fitted <- attributes(
  predict(svmfit.opt, dat[-train, ], decision.values = T)
)$decision.values
rocplot(-fitted, dat[-train, "y"], main = "Test Data")
fitted <- attributes(
  predict(svmfit.flex, dat[-train, ], decision.values = T)
)$decision.values
rocplot(-fitted, dat[-train, "y"], add = T, col = "red")

```

# 9.7 Exercise

## Q5

# 5a

```{r}
x1=runif(500)-0.5
x2=runif(500)-0.5
y=1*(x1^2-x2^2 > 0 )
```

# 5b

```{r}
plot(x1, x2, xlab = "X1", ylab = "X2", col = (4 - y), pch = (3 - y))
```

# 5c

```{r}
df <- data.frame(x1, x2, y = factor(y))
library(glmnet)
glm.fit <-  glm(y~ x1 + x2,family='binomial', data=data.frame(x1,x2,y))
glm.fit
df$pred <- ifelse(predict(glm.fit, type = "response") > 0.5, 1, 0)
summary(glm.fit)
```

# 5d

```{r}
table(df$pred)
glm.pred <- predict(glm.fit,data.frame(x1,x2))
plot(x1,x2,col = ifelse(glm.pred>0,'green','purple'),pch=ifelse(as.integer(glm.pred>0) == y,1,4))
sum(df$y == df$pred) / nrow(df)
```

# 5e

```{r}
glm.fit <- glm(y~ x1 + x2 + I(x1^2) + I(x2^2), family = "binomial", data=data.frame(x1,x2,y))
df$pred <- ifelse(predict(glm.fit, type = "response") > 0.5, 1, 0)
summary(glm.fit)
```

# 5f

```{r}
glm.pred=predict(glm.fit,data.frame(x1,x2))     
plot(x1,x2,col=ifelse(glm.pred>0,'red','darkgrey'),pch=ifelse(as.integer(glm.pred>0) == y,1,4))
```

# 5g

```{r}
library(e1071)
dat <- data.frame(x1 = x1, x2 = x2, y = as.factor(y))
svmfit <- svm(y ~ ., data = dat, kernel = "linear", cost = 10, scale = FALSE)
plot(svmfit, dat)
```

# 5h

```{r}
set.seed(1)
dat <- data.frame(x1 = x1, x2 = x2, y = as.factor(y))
svmfit <- svm(y ~ x1 + x2, data = dat, kernel = "radial", gamma = 1, cost = 1)
plot(svmfit, dat)
svmfit <- svm(y ~ x1 + x2, data = dat, kernel = "radial", gamma = 1, cost = 1e5)
plot(svmfit, dat)
```
# 5i

#The graphs highlight the effectiveness of support vector machines (SVMs) with non-linear kernels in delineating non-linear boundaries

#Logistic regression with non-interactions and SVMs with linear kernels fail to find a clear decision boundary

#Introducing interaction terms to logistic regression appears to empower it similarly to radial-basis kernels

#Though, selecting the appropriate interaction terms involves manual effort and tuning, and this becomes challenging with a substantial number of features

#Radial basis kernels necessitate tuning only parameters gamma and cost facilitating an easier tuning process through cross-validation


## Q8

# 8a

```{r}
library(ISLR)
library(e1071)
set.seed(42)
train=sample(1:1070,800)
test=(1:1070)[-train]
n_train <- 800
n_test = length(test)
```

# 8b

```{r}
svm.fit=svm(Purchase~.,data=OJ,subset=train,cost=0.01,kernel='linear', scale = TRUE)
summary(svm.fit)
```

# 8c

```{r}
svm.pred <- predict(svm.fit, newdata = OJ[train, ])
table(predicted = svm.pred, truth = OJ[train, ]$Purchase)
mean(OJ$Purchase[train] != svm.pred)
svm.pred <- predict(svm.fit, newdata = OJ[test, ])
table(predicted = svm.pred, truth = OJ[test, ]$Purchase)
mean(OJ$Purchase[test] != svm.pred)
```

# 8d

```{r}
svm.tune=tune(svm, Purchase~.,data=OJ[train,], ranges=data.frame(cost=seq(0.01,10,25)), kernel='linear')
summary(svm.tune)
svm.tune=tune(svm, Purchase~.,data=OJ[test,], ranges=data.frame(cost=seq(0.01,10,25)), kernel='linear')
summary(svm.tune)
```

# 8e

```{r}
svm.pred=predict(svm.tune$best.model,OJ[train, ])
table(predicted = svm.pred, truth = OJ[train, ]$Purchase)
train.error = 1 - sum(svm.pred == OJ[train, ]$Purchase)/n_train
train.error
svm.pred=predict(svm.tune$best.model,OJ[test, ])
table(predicted = svm.pred, truth = OJ[test, ]$Purchase)
test.error = 1 - sum(svm.pred == OJ[test, ]$Purchase)/n_test
test.error
```

# 8f

```{r}
svm.fit=svm(Purchase~.,data=OJ,subset=train,cost=0.01,kernel='radial')
summary(svm.fit)
svm.pred=predict(svm.fit,OJ[train,])
table(OJ[train,'Purchase'],svm.pred)
mean(OJ$Purchase[train] != svm.pred)
svm.pred=predict(svm.fit,OJ[test,])
table(OJ[test,'Purchase'],svm.pred)
mean(OJ$Purchase[test] != svm.pred)
svm.tune=tune(svm, Purchase~.,data=OJ[train,], ranges=data.frame(cost=seq(0.01,10,25)), kernel='radial')
summary(svm.tune)
svm.tune=tune(svm, Purchase~.,data=OJ[test,], ranges=data.frame(cost=seq(0.01,10,25)), kernel='radial')
summary(svm.tune)
svm.pred=predict(svm.tune$best.model,OJ[train,])
table(OJ[train,'Purchase'],svm.pred)
mean(OJ$Purchase[train] != svm.pred)
svm.pred=predict(svm.tune$best.model,OJ[test,])
table(OJ[test,'Purchase'],svm.pred)
mean(OJ$Purchase[test] != svm.pred)
```

# 8g:

```{r}
svm.fit=svm(Purchase~.,data=OJ,subset=train,cost=0.01,kernel='polynomial', degree = 2)
summary(svm.fit)
svm.pred=predict(svm.fit,OJ[train,])
table(OJ[train,'Purchase'],svm.pred)
mean(OJ$Purchase[train] != svm.pred)
svm.pred=predict(svm.fit,OJ[test,])
table(OJ[test,'Purchase'],svm.pred)
mean(OJ$Purchase[test] != svm.pred)
svm.tune=tune(svm, Purchase~.,data=OJ[train,], ranges=data.frame(cost=seq(0.01,10,25)), kernel='polynomial', degree = 2)
summary(svm.tune)
svm.pred=predict(svm.tune$best.model,OJ[train,])
table(OJ[train,'Purchase'],svm.pred)
mean(OJ$Purchase[train] != svm.pred)
svm.pred=predict(svm.tune$best.model,OJ[test,])
table(OJ[test,'Purchase'],svm.pred)
mean(OJ$Purchase[test] != svm.pred)


svm.fit=svm(Purchase~.,data=OJ,subset=train,cost=0.01,kernel='polynomial')
summary(svm.fit)
svm.pred=predict(svm.fit,OJ[train,])
table(OJ[train,'Purchase'],svm.pred)
mean(OJ$Purchase[train] != svm.pred)
svm.pred=predict(svm.fit,OJ[test,])
table(OJ[test,'Purchase'],svm.pred)
mean(OJ$Purchase[test] != svm.pred)
svm.tune=tune(svm, Purchase~.,data=OJ[train,], ranges=data.frame(cost=seq(0.01,10,25)), kernel='polynomial')
summary(svm.tune)
svm.pred=predict(svm.tune$best.model,OJ[train,])
table(OJ[train,'Purchase'],svm.pred)
mean(OJ$Purchase[train] != svm.pred)
svm.pred=predict(svm.tune$best.model,OJ[test,])
table(OJ[test,'Purchase'],svm.pred)
mean(OJ$Purchase[test] != svm.pred)
```
# 8h

#Train error for Linear Kernel SVM is 0.165

#Test error for Linear Kernel SVM is 0.18518

#Error estimation of Linear Kernel SVM using 10-fold cross validation: 0.17375

#Train error for Linear Kernel SVM after tuning for optimal cost is 0.165

#Test error for Linear Kernel SVM after tuning for optimal cost is 0.20373


#Train error for Radial Kernel SVM is 0.3975

#Test error for Radial Kernel SVM is 0.3666667

#Error estimation of Radial Kernel SVM using 10-fold cross validation: 0.3975

#Train error for Radial Kernel SVM after tuning for optimal cost is 0.3975

#Test error for Radial Kernel SVM after tuning for optimal cost is 0.3666667


#Train error for Polynimal Kernel with degree 2 SVM is 0.3975

#Test error for Polynimal Kernel with degree 2 SVM is 0.3666667

#Error estimation of Polynimal Kernel with degree 2 SVM using 10-fold cross validation: 0.3975

#Train error for Polynimal Kernel with degree 2 SVM after tuning for optimal cost is 0.3975

#Test error for Polynimal Kernel with degree 2 SVM after tuning for optimal cost is 0.3666667

#Train error and Test error are same for Radial kernel and polynomial kernel with degree 2

#But when the degree for polynomial is not defined, we get different values for Train error and Test error

#Overall the linear SVM gives lowest error rates and hence performs the best
